Architectures for workflow integration in science gateways

### 2016-04-18 paper.tex
Edits in section 5 and 4

The architectures described in Section~\ref{sec:architectures} are
evaluated on Table~\ref{table:evaluation}. We use 5 main criteria to
evaluate the architectures: integration effort, robustness of workflow
execution, extensibility, scalability and other features. Criteria
break down to specific metrics where \emph{low value indicates good
  performance}. For each criterion, a total score is computed by
summing up the individual metrics. We ensure that the different
metrics in a criterion measure similar entities so that summation
makes sense. Metric values obtained by each architecture are
normalized between 0 (best value) and 1 (worst value) to determine the
color of the corresponding table cell (brighter color indicate better
performance\footnote{Normalized metric value $m'$ is
  $\frac{m-m_{\mathrm{min}}}{m_{\mathrm{min}}-m_{\mathrm{max}}}$ where
  $m$ is the initial metric value,
  $m_{\mathrm{min}}$/$m_{\mathrm{max}}$ are the minimal/maximal values
  among all architectures. Cell color is \todo{\#99XX99, where
    X=F-4m'})}. The criteria and metrics are explained hereafter.

@@

\todo{Explain these metrics better, including why the related features
  are useful.}

\emph{Integration} measures the development effort required to build
the architecture. It is obtained by counting the number of
interactions and components on the architectural diagram. It breaks
down to the following 2 metrics:
\begin{itemize}[itemsep=0cm]
\item Components \texttt{I$_1$}),
\item Interactions (\texttt{I$_2$}).
\end{itemize}
One may wonder whether infrastructure should be part of
\texttt{I$_1$}) since it is usually not developed by the groups who
integrate workflow engines in science gateway. However, integrating an
infrastructure in the architecture usually requires some technical
effort (e.g. account creation, software installation, etc) which is
why we keep it in the metric. The two metrics are summed to obtain
the total score for this criterion, which measures the total number of
software pieces to develop.

\emph{Robustness of workflow execution} measures the likelihood that
workflow execution fails due to errors coming from the software
architecture. It is determined as the number of software components and
interactions that are specific to workflow execution. It assume that
complex architectures tend to be prone to failure.  These components
and interactions are represented in red on the architecture diagram in
Figure~\ref{fig:architectures}. Robustness breaks down to 2 metrics:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Components (\texttt{R$_1$}): number of specific components
  involved in workflow executions.
\item Interactions (\texttt{R$_2$}): number of specific interactions
  involved in workflow executions.
\end{itemize}
The two metrics are summed to obtain the total score for this
criterion, which measures the total number of software pieces that are
specifically involved in workflow execution. Robustness of workflow
execution is critical since it directly impacts the end users of the
science gateway.

\emph{Extensibility} measures the difficulty to replace or add
elements in the architecture. It is obtained as the number of
interaction and component to modify when a new element is
added. Modification of an interaction is deemed necessary when the
parameters involved in this interaction are modified. Modification of
a component is required when its code needs to be modified or
recompiled (science gateway or workflow service), or when a new piece
of software has to be installed (infrastructure only). Extensibility
breaks down in 4 metrics depending on the type of element that has to
be added or replaced:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item New engine type (\texttt{E$_1$}): number of interactions or
  components to modify to integrate a new workflow engine type in the
  architecture. Workflow engine belong to different types when they
  cannot be invoked using the same interface. Adding a new type of
  workflow engine allows to execute more workflows in the science gateway.
\item New engine version (\texttt{E$_2$}): number of interactions or
  components to modify to integrate a new version of a workflow engine
  in the architecture, assuming that another version of the same
  engine type is already available. Different versions of a workflow
  engine share the same interface, i.e. they can be invoked using the
  same software. When this is not the case, the different versions
  have to be considered as different engine types. 
\item New workflow (\texttt{E$_3$}): number of interactions required
  to integrate a new workflow in the architecture, assuming that the
  engine type and version required to execute this workflow are
  already available.
\item New infrastructure (\texttt{E$_4$}): number of interactions or components to
  modify to integrate a new type of infrastructure in the
  architecture. Adding a new infrastructure allows to provide more
  computing or storage power, to access specific types of resources
  (e.g. GPUs, clouds), or to enforce execution policies
  (e.g. constrain data to remain in a particular network domain).
\item New science gateway: \todo{Not sure if this makes sense}.
\end{itemize}
The 4 metrics are summed to obtain a global index that measures the
difficulty to extend the architecture.

@@

\caption{Architecture evaluation. Lower values (brighter colors) indicate better performance.}
\label{table:evaluation}
\end{table*}

@@

>>>>

The architectures described in Section~\ref{sec:architectures} are
evaluated in Table~\ref{table:evaluation}. We use 5 main criteria to
evaluate the architectures: integration effort, robustness of workflow
execution, extensibility, scalability and other features. Criteria
break down to specific metrics where \emph{low value indicates good
  performance}. For each criterion, a total score is computed by
summing up the individual metrics. We ensure that the different
metrics in a criterion measure similar entities so that summation
makes sense. The
criteria and metrics are explained hereafter.

@@

\emph{Integration} measures the development effort required to build
the architecture. It is obtained by counting the total number of
interactions and components on the architectural diagram. It breaks
down to the folloowing 2 metrics:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Total number of components (\texttt{I$_1$}),
\item Total number of interactions (\texttt{I$_2$}).
\end{itemize}
One may wonder whether infrastructure should be counted in
\texttt{I$_1$} since it is usually not developed by the groups who
integrate workflow engines in science gateway. However, integrating an
infrastructure in the architecture usually requires some technical
effort (e.g. account creation, software installation, etc) which is
why we keep it in the metric. The two metrics are summed to obtain the
total score for this criterion, which measures the total number of
software pieces to develop.

\emph{Robustness of workflow execution} measures the likelihood that
workflow execution fails due to errors in the components or
interactions of the software architecture. It is a critical criterion
since it directly impacts the end users of the science gateway.
Robustness is determined as the number of software components and
interactions that are specific to workflow execution, i.e. not reused
by other functions:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Components (\texttt{R$_1$}): number of specific components
  involved in workflow executions. Science
gateway, for instance, is \emph{not} specific to workflow execution
since it is used to authenticate users, add new workflows, transfer
user data, etc. 
\item Interactions (\texttt{R$_2$}): number of specific interactions
  involved in workflow executions.
\end{itemize}
These specific components and interactions are represented in red on
the architecture diagrams in Figure~\ref{fig:architectures}. The two
metrics assume that complex architectures tend to be prone to
failure. The two metrics are summed to obtain the total score for this
criterion, which measures the total number of software pieces that are
specifically involved in workflow execution.

\emph{Extensibility} measures the difficulty to replace or add
elements in the architecture. It is determined as the number of
interaction and component to modify when a new element is
added. Modification of an interaction is deemed necessary when the
parameters involved in this interaction are modified. Modification of
a component is required when its code needs to be modified or
recompiled (science gateway or workflow service), or when a new piece
of software has to be installed (infrastructure only). Extensibility
breaks down in 4 metrics depending on the type of element that has to
be added or replaced:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item New engine type (\texttt{E$_1$}): number of interactions or
  components to modify to integrate a new workflow engine type in the
  architecture. Workflow engine belong to different types when they
  cannot be invoked using the same interface. Adding a new type of
  workflow engine allows to execute more workflows in the science gateway.
\item New engine version (\texttt{E$_2$}): number of interactions or
  components to modify to integrate a new version of a workflow engine
  in the architecture, assuming that another version of the same
  engine type is already available. Different versions of a workflow
  engine share the same interface, i.e. they can be invoked using the
  same software. When this is not the case, the different versions
  have to be considered as different engine types. 
\item New workflow (\texttt{E$_3$}): number of interactions required
  to integrate a new workflow in the architecture, assuming that the
  engine type and version required to execute this workflow are
  already available.
\item New infrastructure (\texttt{E$_4$}): number of interactions or components to
  modify to integrate a new type of infrastructure in the
  architecture. Adding a new infrastructure allows to provide more
  computing or storage power, to access specific types of resources
  (e.g. GPUs, clouds), or to enforce execution policies
  (e.g. constrain data to remain in a particular network domain).
\item New science gateway: \todo{Not sure if this makes sense}.
\end{itemize}
The 4 metrics are summed to obtain a global index that measures the
difficulty to extend the architecture.

@@

\caption{Architecture evaluation. Lower values (brighter colors) indicate better performance. On each row, metric values are
  normalized between 0 (best value) and 1 (worst value) to determine the
  color of the corresponding table cell. More precisely, the normalized metric value $m'$ is
  defined as
  $\frac{m-m_{\mathrm{min}}}{m_{\mathrm{min}}-m_{\mathrm{max}}}$ where
  $m$ is the initial metric value, $m_{\mathrm{min}}$/$m_{\mathrm{max}}$
  are the minimal/maximal values among all architectures. The RGB hexadecimal color code of the cell
  is \texttt{\#99XX99}, where X=F-4m'.}
\label{table:evaluation}
\end{table*}

@@

Overall comparison, see Table~\ref{table:overall}.

\begin{table*}
\centering
\begin{tabular}{cccccc}
  \textbf{Tight}
  & \textbf{Service}
  & \textbf{Sub-task}
  & \textbf{Pool}
  & \textbf{Nested}
  & \textbf{Import} \\
\cellcolor[HTML]{99E599}2.4
  & \cellcolor[HTML]{99E599}2.5
  & \cellcolor[HTML]{99FF99}1.5
  & \cellcolor[HTML]{99EF99}2.1
  & \cellcolor[HTML]{99BB99}4.0
  & \cellcolor[HTML]{99FA99}1.7\\
\end{tabular}
\caption{Overall evaluation. Brighter colors indicate better performance.}
\label{table:overall}
\end{table*}
